\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-21 trunk/europa-basics.tex
post-telecon updates to title europa sections

\subsection{The basics}
\label{sec:basics}

@@

\begin{quotation}

  In the near future, a personal robot sets out to buy a gallon of milk
  This involves a number of tasks: obtain keys, obtain wallet,
  start car, drive to store, find and obtain milk, purchase milk, etc.
  The embedded planner has to have a ``model'' of the world in which it
  lives and has to use the task primitives in this model to structure
  the actions so it achieves its goal. Constraints control when certain
  tasks can or cannot occur. For example the robot must obtain the keys
  and wallet \emph{before} driving to the store and pick up the milk
  \emph{before} purchasing it.

\end{quotation}

For such a robot the milk buying plan at the store might look like
\comment{Figure needed}.

>>>>

\paragraph {A brief note on \eu history and notation} The earliest
origin of what we will call \eu in this chapter, is from the
\texttt{HSTS} planner \cite{mus94} which was used for a number of
demonstrations for space telescopes as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was the
\texttt{LISP}-based\footnote{This was the first (and only to our
  knowledge) software ever to been written in \texttt{LISP} to be
  flown in space.} planner that was flown as part of the Remote Agent
\texttt{RAX} experiment. Concurrent to \texttt{RAX}, a C++ version of
the planner was implemented at NASA Ames which went on to be used for
the \texttt{MAPGEN} \cite{bresina05} and continues to be used to date,
in the mission-critical uplink process for the command/control of the
Mars Exploration Rovers (MER) mission. A vastly re-factored, higher
performance open source version \cite{europapso} was named \eut and is
the subject of this chapter. Note however, in using the name \eu, we
refer to the \emph{fundamental ideas as well as the planning paradigm}
even if there exists a specific implementation.

\subsection{The Basics}
\label{sec:basics}

@@

{\small
  \begin{quote}
    In the near future, a personal robot sets out to buy a gallon of
    milk This involves a number of tasks: obtain keys, obtain wallet,
    start car, drive to store, find and obtain milk, purchase milk,
    etc.  The embedded planner has to have a ``model'' of the world in
    which it lives and has to use the task primitives in this model to
    structure the actions so it achieves its goal. Constraints control
    when certain tasks can or cannot occur. For example the robot must
    obtain the keys and wallet \emph{before} driving to the store and
    pick up the milk \emph{before} purchasing it.
\end{quote}

For such a robot the milk buying plan at the store might look like
that shown in Fig. \ref{fig:shop:shopping}.

### 2012-02-21 trunk/europa.tex
post-telecon updates to title europa sections

The \textbf {Language Modules}: Describing problem domains and problem
instances in terms of objects, variables and domain rules in a
programmatic way is cumbersome to the extent of being impractical,
this motivated the creation of modeling languages that support those
definitions in a much more natural and compact way. NDDL is the main
modeling language for \eu and will be described in detail in the
Modeling section below. ANML is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eu,
NDDL (New Domain Description Language [pronounced ``noodle''] allows
the user to specify the different components of a problem domain in a
concise way. The main characteristics that make NDDL a powerful tool
for describing problem domains are:

@@

\item Object Oriented: NDDL supports classes and inheritance in
  similar fashion to popular Object-Oriented languages like Java and
  C++. It also supports polymorphism for some of the planning
  components. Using object classes and instances is a time-tested
  approach to naturally describe problem domains \comment{OO
    citation}.

@@

To understand how the most important elements of NDDL work, we show
how the Shopping Agent problem could be modeled:

@@

\item \textbf{Object Types}: NDDL supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. NDDL supports a single rooted class hierarchy so
  that state and behavior can be extended through inheritance. We took
  advantage of that by making \textit{AgentLocation} inherit from
  \eu's built-in \textit{Timeline} class which doesn't allow
  predicates to overlap in time for the same object instance, in this
  way we can ensure that and agent can only be at one place at a time.

@@

Domain rules can be defined in NDDL in the context of a Predicate or
an Action. In the example above we can see the main elements of a rule
definition:

\begin{enumerate}
\item \textbf{Variables}: A Token Type (predicate or action) is part
  of a class, therefore class attributes (see references to
  \texttt{object.location} above), Token Type parameters (see
  references to \texttt{[from, to]} in \texttt{Agent::Go}, or to
  \texttt{[duration, product]} in \texttt{Agent::Buy} and locally
  declared variables (see \texttt{possibleStores} in
  \texttt{Agent::Buy}) can all be used in the definition of a Rule.

@@

  \item \textbf{Subgoals}: In \texttt{Agent::Go}, subgoals are used to
    specify that the Shopping Agent's location must transition from
    origin to destination in a temporally consistent way. The
    \texttt{[meets, equals, metby]} operators are used to specify
    subgoals (Allen operators are explained in more detail below),
    along with the temporal relationship each token representing a
    subgoal has with the the token representing the \texttt{Agent::Go}
    action. Also, \texttt{[condition, effect]} annotations can be used
    when defining a subgoal inside an action, \eu doesn't immediately
    associate any semantics with them, but they can be extremely
    useful for search modules when creating plans.
\end{enumerate}

@@

In \eu,  all the Variables and Constraints from a problem domain domain description are kept inside the Constraint Reasoning Engine (CRE).
The CRE provides a framework where any number of constraint types can collaborate to perform inference and prune variable domains using Bounds Propagation and Domain Reduction techniques. \eu uses a Propagator pattern, in which a Propagator object keeps track of a set of constraint instances of the same type and is responsible for ensuring that whenever any variable changes or constraints are introduced or dropped, all related variables are updated through inference.

@@

\comment {add europa's AC-3 algorithm here}

@@

\textit{Temporal Constraints:} \eu uses variables to explicitly represent timepoints for
plan activities and states. Constraints among timepoints provide a
natural way to express domain axioms. For example, in order to state
that activity A must occur before activity B we can say that the end
timepoint of A is $leq$ the start timepoint of B.

@@

\item \textbf {Unary} resources are Reusable resources with capacity=1, which allows \eu to perform 
some run time savings compared to the generic Reusable class.
\end{enumerate}

@@

Now we have all the elements in place so that an automated problem
solver can be created. Let's recap what those elements are:

@@

\item A problem instance (also called initial state) that consists of:
    \begin{enumerate}
        \item Variable and Object instances that exist for the entire planning
  horizon.
        \item Temporally scoped predicate and action instances.
        \item Temporally scoped goals.
\end{enumerate}

@@

\begin{enumerate}
\item Find the flaws of PP, that is the conditions that prevent it
  from being a solution plan (flaw types are explained below).
\item Select one such flaw
\item Select a resolver for the flaw
\item Refine PP by applying the resolver
\item If an inconsistency is found, try another resolver.
\item If all possible resolvers for a particular flaw fail, return
  failure, otherwise continue until resolving all the flaws.
\end{enumerate}

This Plan Space Planning algorithm is translated into \eu's
representation as follows:

@@

\item Line \ref{li:noplan}: If we arrive here, then we have exhausted
  all options to resolve the flaw, including the case where no options
  were available initially. Thus the problem cannot be solved.
\end{enumerate}

@@

\eu's Constraint Reasoning Engine can be used to compute the non-attack constraint violations (\texttt{ce.getViolations()} is a straight call on the \eu API), 
to easily pick the queen that has he most violations by implementing \texttt{getQueenWithMaxViolation(ce)} and to look ahead and see what 
moves are possible for a queen that would remove its violations by implementing \texttt{getMoves(ce,queenToMove,curPos)}.

>>>>

The \textbf {Language Modules}: Describing problem domains and problem
instances in terms of objects, variables and domain rules in a
programmatic way is cumbersome to the extent of being impractical,
this motivated the creation of modeling languages that support those
definitions in a much more natural and compact way. \texttt{NDDL} (New
Domain Description Language [pronounced ``noodle'']) \cite{NDDL} is
the main modeling language for \eu and will be described in detail in
the Modeling section below. ANML is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eu,
\texttt{NDDL} allows the user to specify the different components of a
problem domain in a concise way. The main characteristics that make
\texttt{NDDL} a powerful tool for describing problem domains are:

@@

\item Object Oriented: \texttt{NDDL} supports classes and inheritance
  in similar fashion to popular Object-Oriented languages like Java
  and C++. It also supports polymorphism for some of the planning
  components. Using object classes and instances is a time-tested
  approach to naturally describe problem domains \cite{rumbaugh}.

@@

To understand how the most important elements of \texttt{NDDL} work,
we show how the Shopping Agent problem could be modeled:

@@

\item \textbf{Object Types}: \texttt{NDDL} supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. \texttt{NDDL} supports a single rooted class
  hierarchy so that state and behavior can be extended through
  inheritance. We took advantage of that by making
  \textit{AgentLocation} inherit from \eu's built-in \textit{Timeline}
  class which doesn't allow predicates to overlap in time for the same
  object instance, in this way we can ensure that and agent can only
  be at one place at a time.

@@

Domain rules can be defined in \texttt{NDDL} in the context of a
Predicate or an Action. In the example above we can see the main
elements of a rule definition:

\item \textbf{Variables}: A Token Type (predicate or action) is part
  of a class, therefore class attributes (see references to
  \texttt{object.location} above), Token Type parameters (see
  references to \texttt{[from, to]} in \texttt{Agent::Go}, or to
  \texttt{[duration, product]} in \texttt{Agent::Buy} and locally
  declared variables (see \texttt{possibleStores} in
  \texttt{Agent::Buy}) can all be used in the definition of a Rule.

@@

\item \textbf{Subgoals}: In \texttt{Agent::Go}, subgoals are used to
  specify that the Shopping Agent's location must transition from
  origin to destination in a temporally consistent way. The
  \texttt{[meets, equals, metby]} operators are used to specify
  subgoals (Allen operators are explained in more detail below), along
  with the temporal relationship each token representing a subgoal has
  with the the token representing the \texttt{Agent::Go} action. Also,
  \texttt{[condition, effect]} annotations can be used when defining a
  subgoal inside an action, \eu doesn't immediately associate any
  semantics with them, but they can be extremely useful for search
  modules when creating plans.

@@

In \eu, all the Variables and Constraints from a problem domain,
domain descriptions are kept inside the Constraint Reasoning Engine
(CRE).  The CRE provides a framework where any number of constraint
types can collaborate to perform inference and prune variable domains
using Bounds Propagation and Domain Reduction techniques. \eu uses a
Propagator pattern, in which a Propagator object keeps track of a set
of constraint instances of the same type and is responsible for
ensuring that whenever any variable changes or constraints are
introduced or dropped, all related variables are updated through
inference.

@@

\kcomment{add europa's AC-3 algorithm here}

@@

\textit{Temporal Constraints:} \eu uses variables to explicitly
represent timepoints for plan activities and states. Constraints among
timepoints provide a natural way to express domain axioms. For
example, in order to state that activity $A$ must occur before
activity $B$ we can say that the end timepoint of $A$ is $leq$ the
start timepoint of $B$.

@@

\item \textbf {Unary} resources are Reusable resources with
  capacity=1, which allows \eu to perform some run time savings
  compared to the generic Reusable class.

@@

Given the above elements, an automated problem solver can be
created using:

@@

\item A problem instance (also called initial state) that consists of:

    \item Variable and Object instances that exist for the entire
      planning horizon.

    \item Temporally scoped predicate and action instances.

    \item Temporally scoped goals.

@@

\item Find the flaws of PP, that is the conditions that prevent it
  from being a solution plan (flaw types are explained below).

\item Select one such flaw

\item Select a resolver for the flaw

\item Refine PP by applying the resolver

\item If an inconsistency is found, try another resolver.

\item If all possible resolvers for a particular flaw fail, return
  failure, otherwise continue until resolving all the flaws.

This Plan Space Planning algorithm is translated into \eus
representation as follows:

@@

\item Line \ref{li:noplan}: If we arrive here, then we have exhausted
  all options to resolve the flaw, including the case where no options
  were available initially. Thus the problem cannot be solved.

@@

\eus Constraint Reasoning Engine can be used to compute the non-attack
constraint violations (\texttt{ce.getViolations()} is a straight call
on the \eu API), to easily pick the queen that has he most violations
by implementing \texttt{getQueenWithMaxViolation(ce)} and to look
ahead and see what moves are possible for a queen that would remove
its violations by implementing
\texttt{getMoves(ce,queenToMove,curPos)}.
